---
title: "NBCUniversal - Forecasting Model Building and Evaluation"
author: "Chunzi Wang"
date: "March 20, 2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


Predictive Modeling


1. In order to do Linear Regression, we don't want useless columns including title, imdbid, dvd_release, release_date (keep seperate release info) and roi so we only keep the useful numeric columns as predictors.

```{r}
keep <- c("imdb_rating","metacritic","imdb_votes","Budget","Box.Office.Gross","release.year","release.month")
movies.clean <- movies.before[keep]
```

```{r}
head(movies.clean)
```

3. For runtime column we remove the "min" and transform it into numeric type.

```{r}
for (i in 1:nrow(movies.before)){
  movies.clean$runtime[i] <- as.numeric(strsplit(as.character(movies.before$runtime),split=" ")[[i]][1])
}
```

4. For other categorical columns including rating, production, actors, director, genre, awards, keywords, we will use bag of words method to create a dtm and add these dummy variable to train.clean.

```{r}
library(tm)
```

4.1 Genre

```{r}
genre <- as.character(movies.before$genre)
head(genre,5)
```

```{r}
genre.source <- VectorSource(genre)
genre.corpus <- VCorpus(genre.source)
print(genre.corpus[[1]][1])
```

Create a clean_corpus function.

```{r}
clean_corpus <- function(corpus){
  corpus <- tm_map(corpus,removePunctuation)
  return(corpus)
}
```

```{r}
genre.corpus.clean <- clean_corpus(genre.corpus)
print(genre.corpus.clean[[1]][1])
```

```{r}
genre.dtm <- DocumentTermMatrix(genre.corpus.clean)
genre.matrix <- as.matrix(genre.dtm)
dim(genre.matrix)
genre.matrix[1:5,1:23]
```

```{r}
term.freq <- colSums(genre.matrix)
term.freq <- sort(term.freq,decreasing = TRUE)
term.freq
```

```{r}
keep.genre <- c("action","adventure","comedy","drama","romance","fantasy","horror","thriller","mystery","scifi","crime")
genre.df <- as.data.frame(genre.matrix)[keep.genre]
head(genre.df)
```

4.2 Actor

```{r}
# or first name and last name will be seperated into two items
actor <- as.character(movies.before$actors)
actor <- gsub(",", "   ", actor, fixed = TRUE)
actor <- gsub(" ", ".", actor, fixed = TRUE)
actor <- gsub("...", " ", actor, fixed = TRUE)
actor <- gsub(".", "", actor, fixed = TRUE)

actor.source <- VectorSource(actor)
actor.corpus <- VCorpus(actor.source)
actor.corpus.clean <- clean_corpus(actor.corpus)

actor.dtm <- DocumentTermMatrix(actor.corpus.clean)
actor.matrix <- as.matrix(actor.dtm)

term.freq <- colSums(actor.matrix)
term.freq <- sort(term.freq,decreasing = TRUE)
head(term.freq,50)
```

```{r}
keep.actor <- c("robertdeniro","tomhanks","adamsandler","liamneeson","stevecarell","dwaynejohnson","jasonbateman","markwahlberg","naomiwatts","robertdowneyjr","brucewillis","juliannemoore","mattdamon","matthewmcconaughey","owenwilson","woodyharrelson","amyadams","jenniferaniston","johnnydepp","arnoldschwarzenegger","jasonstatham","jeffbridges","markruffalo","nicolascage","susansarandon","alecbaldwin","bradleycooper","channingtatum","danaykroyd","dennisquaid","emilyblunt","ewanmcgregor","georgeclooney","jamesfranco","jamesmcavoy","nicolekidman","camerondiaz","juliaroberts","jenniferlawrence")
actor.df <- as.data.frame(actor.matrix)[keep.actor]
head(actor.df)
```

4.3 Production

```{r}
production <- as.character(movies.before$production)
production <- gsub(" ","",production,fixed=TRUE)
production.source <- VectorSource(production)
production.corpus <- VCorpus(production.source)

production.dtm <- DocumentTermMatrix(production.corpus)
production.matrix <- as.matrix(production.dtm)

term.freq <- colSums(production.matrix)
term.freq <- sort(term.freq,decreasing = TRUE)
head(term.freq,20)
```

```{r}
keep.production <- c("universalpictures","warnerbros.pictures","20thcenturyfox","sonypictures","paramountpictures","waltdisneypictures","focusfeatures","lionsgatefilms","sonypicturesclassics","columbiapictures")
production.df <- as.data.frame(production.matrix)[keep.production]
```

4.4 Director

```{r}
director <- as.character(movies.before$director)
director <- gsub(" ","",director,fixe=TRUE)
director.source <- VectorSource(director)
director.corpus <- VCorpus(director.source)

director.dtm <- DocumentTermMatrix(director.corpus)
director.matrix <- as.matrix(director.dtm)

term.freq <- colSums(director.matrix)
term.freq <- sort(term.freq,decreasing = TRUE)
```
```{r}
head(term.freq,70)
```

```{r}
keep.director <- c("stevenspielberg","jamescameron","christophernolan","woodyallen","ronhoward","robertzemeckis","stevensoderbergh","dennisdugan","spikelee","oliverstone","peterberg","ridleyscott","anglee","robcohen","timburton","jameswan","michaelbay","clinteastwood","tylerperry","peterjackson","ethancoen,joelcoen","davidlynch","joewright","sofiacoppola")
director.df <- as.data.frame(director.matrix)[keep.director]
```

4.5 Award and nominaton

Here the infos about awards and nomination has been reduced to four columns: nomination, oscar, golden globe, and award. Numbers of awards and nominations are not included for simplicity reason.

```{r}
library(textreg)
```

```{r}
award <- as.character(movies.before$awards)
award.source <- VectorSource(award)
award.corpus <- VCorpus(award.source)

# clean and stem corpus
award.corpus <- tm_map(award.corpus,removePunctuation)
award.corpus <- tm_map(award.corpus,removeNumbers)
award.corpus <- tm_map(award.corpus,content_transformer(tolower))
award.corpus <- stem.corpus(award.corpus)

award.dtm <- DocumentTermMatrix(award.corpus)
award.matrix <- as.matrix(award.dtm)

term.freq <- colSums(award.matrix)
term.freq <- sort(term.freq,decreasing = TRUE)
term.freq
```

```{r}
keep.award <- c("nomin+","oscar+","globe+","award+")
award.df <- as.data.frame(award.matrix)[keep.award]
colnames(award.df) <- c("nomination","oscar","golden globe","award")
```

4.6 Keyword

```{r}
keyword <- as.character(movies.before$keywords)
keyword <- gsub("|"," ",keyword,fixe=TRUE)
keyword <- gsub("-"," ",keyword,fixe=TRUE)

keyword.source <- VectorSource(keyword)
keyword.corpus <- VCorpus(keyword.source)
keyword.corpus <- tm_map(keyword.corpus,removeWords,stopwords("en"))

keyword.dtm <- DocumentTermMatrix(keyword.corpus)
keyword.matrix <- as.matrix(keyword.dtm)

term.freq <- colSums(keyword.matrix)
term.freq <- sort(term.freq,decreasing = TRUE)
head(term.freq,100)
```

```{r}
keep.keyword <- c("relationship","death","shot","car","female","woman","man","male","police","sex","nudity","blood","gun","murder","hero","father","mother","son","american","daughter","child","wife","fight","school","husband","violence","love","family","dog","war","animal","friendship","teenage")
keyword.df <- as.data.frame(keyword.matrix)[keep.keyword]
head(keyword.df)
```

4.7 Rating

I don't know why some ratings in corpus cannot be transalted into dtm. Just ignore it for now

```{r}
rating <- as.character(movies.before$rating)
head(rating,10)
rating.source <- VectorSource(rating)
rating.corpus <- VCorpus(rating.source)

table(movies.before$rating)

for (i in 1:20){
  print(rating.corpus[[i]][1])
}
```

```{r}
rating.dtm <- DocumentTermMatrix(rating.corpus)
rating.matrix <- as.matrix(rating.dtm)
dim(rating.matrix)
head(rating.matrix)
```


5. Combine these dataframes together

```{r}
dim(movies.clean)
dim(actor.df)
dim(award.df)
dim(director.df)
dim(genre.df)
dim(keyword.df)
dim(production.df)
```

```{r}
movies.clean.df <- cbind(movies.clean,actor.df,award.df,director.df,genre.df,keyword.df,production.df)
dim(movies.clean.df)
```

```{r}
write.csv(movies.clean.df,"movies_clean.csv")
```

6. split into train and test set for model validation.

Find NAs and in movies.clean.df in imdb_rating, matacritic, imdb_votes, and runtime.

```{r}
summary(movies.clean.df)
```

```{r}
# Fill in 0 where there's NA
movies.clean.df[which(is.na(movies.clean.df$imdb_rating)),1] <- 0
movies.clean.df[which(is.na(movies.clean.df$metacritic)),2] <- 0
movies.clean.df[which(is.na(movies.clean.df$imdb_votes)),3] <- 0
movies.clean.df[which(is.na(movies.clean.df$runtime)),8] <- 0
```

```{r}
set.seed(1)
train.index <- sample(nrow(movies.clean.df),round(nrow(movies.clean.df)*0.8))
train <- movies.clean.df[train.index,]
test <- movies.clean.df[-train.index,]
```

7. Fit linear regression model

```{r}
train.fit <- lm(Box.Office.Gross~.,data=train)
summary(train.fit)
```

In this model, p-value is significant and we have good explanatory power (68.67% of variability in box office is expalined by these predictors.) given the complexity of the dataset.

```{r}
plot(train.fit)
```

We can see that observation 11, 811, 725 tend to be outliers in three plots, so we elimate them from the train set and fit the model again. There's a light improvement but so little that could be ignored. 

Predict using test set and calculate rmse.

```{r}
pred <- predict(train.fit,test)
test$pred.lm <- pred
```

There're negative predicted value here, which means our model is not performing very well.

```{r}
pred.result <- test %>%
  select(Box.Office.Gross,pred.lm)
pred.result$ratio.lm <- round(pred.result$pred.lm/pred.result$Box.Office.Gross,2)
pred.result
```

```{r}
RMSE.lm <- sqrt(mean((test$pred.lm-test$Box.Office.Gross)^2))
RMSE.lm
```

Subset the dataset using only siginificant predictors in lm. It doesn't impact the r-squared at all. Disapointed.

```{r}
keep.predictor <- c("imdb_rating","imdb_votes","stevecarell","dwaynejohnson","robertdowneyjr","owenwilson","arnoldschwarzenegger","bradleycooper","camerondiaz","jenniferlawrence","nomination","oscar","golden globe","award","stevenspielberg","jamescameron","robertzemeckis","ronhoward","jameswan","michaelbay","tylerperry","joewright","adventure","drama","death","female","man","nudity","blood","murder","hero","mother","love","animal","friendship","universalpictures","warnerbros.pictures","20thcenturyfox","sonypictures","waltdisneypictures","columbiapictures","sonypicturesclassics","Box.Office.Gross")
train.new <- train[keep.predictor]
test.new <- test[keep.predictor]
```

```{r}
train.fit <- lm(Box.Office.Gross~.,data=train.new)
summary(train.fit)
```

```{r}
pred <- predict(train.fit,test.new)
test.new$pred.lm <- pred
pred.result <- test.new %>%
  select(Box.Office.Gross,pred.lm)
pred.result$ratio.lm <- round(pred.result$pred.lm/pred.result$Box.Office.Gross,2)
pred.result
```

```{r}
RMSE.lm <- sqrt(mean((test.new$pred.lm-test.new$Box.Office.Gross)^2))
RMSE.lm
```

8. Fit regression tree model

```{r}
library(rpart)
library(rpart.plot)
```

```{r}
tree.fit <- rpart(formula=Box.Office.Gross~.,data=train,method="anova")
print(tree.fit)
rpart.plot(tree.fit)
```

```{r}
pred.tree <- predict(tree.fit,test)

pred.result$pred.tree <- pred.tree
pred.result$ratio.tree <- round(pred.result$pred.tree/pred.result$Box.Office.Gross,2)
pred.result
```

```{r}
RMSE.tree <- sqrt(mean((pred.tree-test$Box.Office.Gross)^2))
RMSE.tree
```

9. Fit random forest model

```{r}
library(randomForest)
```

```{r}
names(train) <- make.names(names(train))
rf.fit <- randomForest(formula=Box.Office.Gross~.,data=train,importance=TRUE)
rf.fit
varImpPlot(rf.fit)
```

70.47% of variance explained, better than linear regression with lower rmse.

```{r}
names(test) <- make.names(names(test))
pred.rf <- predict(rf.fit,test)

pred.result$pred.rf <- pred.rf
pred.result$ratio.rf <- round(pred.rf/pred.result$Box.Office.Gross,2)
pred.result
```

```{r}
RMSE.rf <- sqrt(mean((pred.rf-test$Box.Office.Gross)^2))
RMSE.rf
```

10. Fit xgboost

```{r}
library(xgboost)
```

```{r}
xgb.fit <- xgboost(data=as.matrix(train[,-5]),
                   nfold=5,
                   label=as.matrix(train$Box.Office.Gross),
                   nrounds=2200,
                   verbose=FALSE,
                   objective="reg:linear",
                   eval_metric="rmse",
                   nthread=8,
                   eta=0.01,
                   gamma=0.0468,
                   max_depth=15,
                   min_child_weight=1.7817,
                   subsample=0.5213,
                   colsample_bytree=0.4603)
```

```{r}
xgb.fit
```

```{r}
library(Metrics)
```

```{r}
pred.xgb <- predict(xgb.fit,newdata=as.matrix(test[,-5]))

pred.result$pred.xgb <- pred.xgb
pred.result$ratio.xgb <- round(pred.xgb/pred.result$Box.Office.Gross,2)
pred.result
```

```{r}
RMSE.xgb <- rmse(test$Box.Office.Gross,pred.xgb)
RMSE.xgb
```

xgb has the lowest rmse among the 4 models.









